### 2014-04-25 Response_to_R0_Reviews.tex

    ARFI resolution within a native imaging frame is not considered a limiting
    factor in our ability to delineate the prostate central glands since the
    scan-conversion and slice decimation introduce resolution distortions that
    are at least an order-of-magnitude larger than our native image plane
    resolution.  ARFI image resolution would be a more realistic limiting
    factor if we were using those primary imaging frames for exclusive gland
    delineation, but our use of the 3 orthogonal planes for segmentation makes
    this resolution of higher-order parameter.  We have added to the following
    text to Section V to emphasize this point to the reader:

    \input{arfi_resolution}

    \item \textit{The authors explain that gross pathology weights and axis
            measurements were affected by the presence of peri-prostatic
            tissue.  Why could the peri-prostatic tissue not be differentiated
            from the prostatic tissue in histology after processing?  Wouldn't
            this enable better gold standard axis measurements?}

    The extraprostatic tissue that is left on the prostate by the surgeon is
    usually minimal, but can vary case-by-case and is left in place during
    gross and histologic processing to adhere to AJCC guidelines that require
    the reporting of potential extraprostatic invasion of malignany tumors.
    This extraprostatic tissue is left in place during whole-mount histology
    processing, but given the challenges in estimating prostate gland size in
    histology (detailed in the response to the second reviewer below), our
    pathologist (AB) believed that the most accurate estimations of \invivo
    prostate dimensions would be taken from the gross pathology specimens.  In
    consulting with our pathologist on this specific comment, his impression
    was that the impact of extraprostatic tissue would be minimal,  so we have
    removed this section from the discussion to focus on more relevant points.

@@

\textit{This is a nice study of ARFI imaging of the prostate that compares ARFI images
to MRI. Three areas of the paper could be improved:}

\begin{itemize}
    \item \textit{The ellipsoidal prostate volume estimate from pathology could be
        improved by using the prostate segmentation on each of the whole mount
        pathology slides to align the data to the MRI prostate surface. The
        volume comparison is further complicated by the shrinkage of the
        prostate after fixation.}

   Our ellipsoidal approximation of the prostate volume is a rough
   approximation from the surgical pathology reports, but we felt it was our
   best metric to approximate prostate size for the following reasons:

   \begin{itemize}
        \item The gross surgical pathology measurements used to make the
            ellipsoidal volume approximation were done after peri-prostatic
            tissue was dissected away, but before formalin fixation, avoiding
            the complication of tissue shrinkage associated with formalin
            fixation, which our pathologists estimated could be as much as 25\%
            of the prostate volume (citation provided in new text).
        \item Whole mount histology slides were explored for this purpose, but
            they suffer from post-fixation contraction, $\sim$ 3 mm slice
            thickness, and slice tearing and folding that compromise accurate
            cross-sectional area representations on a single slide.
            Additionally, the whole mount histology slides require some degree
            of image registration due to their arbitrary orientation on the
            slides, adding another source of error.
    \end{itemize}

### 2014-04-25 abstract.tex
updated response on extraprostatic tissue based on Buck feedback

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second-leading cause of cancer-related death.
Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
popularity to characterize PCa.  ARFI imaging has the potential to aid PCa
diagnosis and management by using tissue stiffness to evaluate prostate zonal
anatomy and lesions.  MR and \textbf{B-mode/}ARFI \invivo imaging datasets were
compared to one another and with gross pathology measurements made immediately
after radical prostatectomy.  Images were manually segmented in 3D Slicer to
delineate the central gland (CG) and prostate capsule, and 3D models were
rendered to evaluate zonal anatomy dimensions and volumes.  Both imaging
modalities showed good correlation between estimated organ volume and gross
pathologic weights.  Ultrasound and MR total prostate volumes were
well-correlated (R$^2$ = \MRarfiVolTotalRsq), but \textbf{B-mode} images
yielded prostate volumes that were larger (\MRarfiVolTotalMeanDiff\% $\pm$
\MRarfiVolTotalStdDiff\%) than MR images, due to overestimation of the lateral
dimension (\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), with
less significant differences in the other dimensions
(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$
\ARFImrTotalApexBaseStdPct\%, apex-to-base).  ARFI and MR CG volumes were also
well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
attributed to ARFI underestimation of the apex-to-base axis
(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and ARFI
overestimation of the lateral dimension (\ARFImrCentralLatLatMeanPct $\pm$
\ARFImrCentralLatLatStdPct\%).  B-mode/ARFI imaging yielded prostate volumes
and dimensions that were well-correlated with MR T2WI estimates, with biases in
the lateral dimension due to poor contrast caused by peri-prostatic fat.
\textbf{B-mode combined with} ARFI imaging is a promising low-cost, portable,
real-time modality that can complement mpMRI for PCa diagnosis, treatment
planning and management.

>>>>

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second-leading cause of cancer-related death.
Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
popularity to characterize PCa.  ARFI imaging has the potential to aid PCa
diagnosis and management by using tissue stiffness to evaluate prostate zonal
anatomy and lesions.  MR and \textbf{B-mode/}ARFI \invivo imaging datasets were
compared to one another and with gross pathology measurements made immediately
after radical prostatectomy.  Images were manually segmented in 3D Slicer to
delineate the central gland (CG) and prostate capsule, and 3D models were
rendered to evaluate zonal anatomy dimensions and volumes.  Both imaging
modalities showed good correlation between estimated organ volume and gross
pathologic weights.  Ultrasound and MR total prostate volumes were
well-correlated (R$^2$ = \MRarfiVolTotalRsq), but \textbf{B-mode} images
yielded prostate volumes that were larger (\MRarfiVolTotalMeanDiff\% $\pm$
\MRarfiVolTotalStdDiff\%) than MR images, due to overestimation of the lateral
dimension (\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%), with
less significant differences in the other dimensions
(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$
\ARFImrTotalApexBaseStdPct\%, apex-to-base).  ARFI and MR CG volumes were also
well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
attributed to ARFI underestimation of the apex-to-base axis
(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and ARFI
overestimation of the lateral dimension (\ARFImrCentralLatLatMeanPct $\pm$
\ARFImrCentralLatLatStdPct\%).  B-mode/ARFI imaging yielded prostate volumes
and dimensions that were well-correlated with MR T2WI estimates, with biases in
the lateral dimension due to poor contrast caused by extraprostatic fat.
\textbf{B-mode combined with} ARFI imaging is a promising low-cost, portable,
real-time modality that can complement mpMRI for PCa diagnosis, treatment
planning and management.

### 2014-04-25 disc_arfi_mr.tex
updated response on extraprostatic tissue based on Buck feedback

The lateral boundaries of the prostate in the \textbf{ultrasound} dataset were
overestimated relative to MR T2WI, increasing \textbf{B-mode} total volume
relative to MR. The lateral overestimation was a result of poor contrast
between the PZ and peri-prostatic fat in the B-mode images, and image
decimation along the lateral dimension in both ARFI and B-mode images. Without
image decimation, the elevation and axial orthogonal planes can be used to
correct errors introduced by the poor contrast between the PZ and
peri-prostatic fat. However, with image decimation, segmentation changes in the
elevation and axial orthogonal planes were limited to the lower bound on the
model precision of, again, $\sim$3 mm.  While reduced decimation could improve
the precision of the lateral measurements, this would have introduced
significant overhead in the manual segmentation process, with diminishing
returns given the spatial smoothing of the modeling algorithms that were
utilized.

>>>>

The lateral boundaries of the prostate in the \textbf{ultrasound} dataset were
overestimated relative to MR T2WI, increasing \textbf{B-mode} total volume
relative to MR. The lateral overestimation was a result of poor contrast
between the PZ and extraprostatic fat in the B-mode images, and image
decimation along the lateral dimension in both ARFI and B-mode images. Without
image decimation, the elevation and axial orthogonal planes can be used to
correct errors introduced by the poor contrast between the PZ and
extraprostatic fat. However, with image decimation, segmentation changes in the
elevation and axial orthogonal planes were limited to the lower bound on the
model precision of, again, $\sim$3 mm.  While reduced decimation could improve
the precision of the lateral measurements, this would have introduced
significant overhead in the manual segmentation process, with diminishing
returns given the spatial smoothing of the modeling algorithms that were
utilized.

### 2014-04-25 image_proc_methods.tex
updated response on extraprostatic tissue based on Buck feedback

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented by a \textbf{single radiologist
    (CK)} using the smooth polygon tool in ITK-SNAP
(http://www.itksnap.org)~\cite{Yushkevich2006}, using unique labels for the PZ,
CG and AFS. The gland was segmented from base-to-apex.  The base was identified
below the bladder, and subsequent images were segmented until the last slice
with visible prostatic tissue was identified caudally. The CG, urethra, PZ and
AFS were segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby peri-prostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.  \textbf{All MR image segmentations were
    ultimately reviewed by an experienced prostate radiologist (RTG).}

@@

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
\textbf{by two researchers (TJG \& ZAM)} using the \verb+draweffect+ tool to
model and measure the prostate capsule and CG volume and tri-axial dimensions
(Figure~\ref{fig:arfi_segs}).  \input{bmode_arfi_coregistered} The sagittal imaging plane was chosen for
segmentation because it has a greater number of easily identifiable anatomical
structures:  seminal vesicles in the left and right posterior base, urethra in
the mid apex, surrounding fibromuscular tissue in the anterior, and rectum in
the posterior. Prostate capsule was segmented in B-mode as this modality offers
both high contrast between prostatic tissue and surrounding peri-prostatic fat,
and better SNR in the anterior capsule region than ARFI images. ARFI images
were used for CG segmentation.  Prior to segmentation, both the B-mode and ARFI
image stacks were down-sampled along the sagittal axis to a slice thickness of
3 mm to reduce the number of manually-segmented frames (18 vs. 360 slices /
image stack) and match the MRI slice thickness.

>>>>

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented by a \textbf{single radiologist
    (CK)} using the smooth polygon tool in ITK-SNAP
(http://www.itksnap.org)~\cite{Yushkevich2006}, using unique labels for the PZ,
CG and AFS. The gland was segmented from base-to-apex.  The base was identified
below the bladder, and subsequent images were segmented until the last slice
with visible prostatic tissue was identified caudally. The CG, urethra, PZ and
AFS were segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby extraprostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.  \textbf{All MR image segmentations were
    ultimately reviewed by an experienced prostate radiologist (RTG).}

@@

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
\textbf{by two researchers (TJG \& ZAM)} using the \verb+draweffect+ tool to
model and measure the prostate capsule and CG volume and tri-axial dimensions
(Figure~\ref{fig:arfi_segs}).  \input{bmode_arfi_coregistered} The sagittal imaging plane was chosen for
segmentation because it has a greater number of easily identifiable anatomical
structures:  seminal vesicles in the left and right posterior base, urethra in
the mid apex, surrounding fibromuscular tissue in the anterior, and rectum in
the posterior. Prostate capsule was segmented in B-mode as this modality offers
both high contrast between prostatic tissue and surrounding extraprostatic fat,
and better SNR in the anterior capsule region than ARFI images. ARFI images
were used for CG segmentation.  Prior to segmentation, both the B-mode and ARFI
image stacks were down-sampled along the sagittal axis to a slice thickness of
3 mm to reduce the number of manually-segmented frames (18 vs. 360 slices /
image stack) and match the MRI slice thickness.

### 2014-04-25 limitations.tex
updated response on extraprostatic tissue based on Buck feedback

\subsection{Limitations}
This study has several limitations that should be considered when interpreting
these results.  Gross pathology weight and axis measurements could both be
affected by the presence of peri-prostatic tissue that was excised during
radical prostatectomy, especially in cases where more aggressive margins may
have been necessary.  For this reason, unlike the image-to-image measurement
comparisons, all of the image metrics presented relative to pathology metrics
(Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
but instead, relative correlations were evaluated.  Additionally, the volumes
of the prostate from gross pathologic measurements were approximated as
ellipsoids, which also introduced error, most likely an over-estimation of
volume.  Interestingly, all pathology estimates were thought to have positive
biases, but both imaging modalities tended to overestimate volume relative to
the pathology measurements.

>>>>

\subsection{Limitations}
This study has several limitations that should be considered when interpreting
these results.  \st{Gross pathology weight and axis measurements could both be
affected by the presence of extraprostatic tissue that was excised during
radical prostatectomy, especially in cases where more aggressive margins may
have been necessary.  For this reason, unlike the image-to-image measurement
comparisons, all of the image metrics presented relative to pathology metrics
(Figure~\ref{fig:mr_arfi_weight}) were not characterized for absolute accuracy,
but instead, relative correlations were evaluated.}  Additionally, the volumes
of the prostate from gross pathologic measurements were approximated as
ellipsoids, which also introduced error, most likely an over-estimation of
volume.  Interestingly, all pathology estimates were thought to have positive
biases, but both imaging modalities tended to overestimate volume relative to
the pathology measurements.
